---
title: "Famous for Forever: an Analysis on Coldplay and Metallica Data"
author: "Talyah Greyling"
date: "`r Sys.Date()`"
bibliography: Tex/ref.bib       
output:
  pagedown::html_paged:
    # template: wp_paged.html
    # css: ['wp.css', 'wp-fonts.css', 'wp-page.css']
    css: ["Template/default-fonts-Texevier.css", "Template/default-page-Texevier.css", "Template/default-Texevier.css"]
    csl: Template/harvard-stellenbosch-university.csl # referencing format used.
    template: ["Template/paged-Texevier.html"]

    toc: true 
    # change to true for a self-contained document, but it'll be a litte slower for Pandoc to render
    self_contained: TRUE
    
abstract: |
    This report contains an analysis on Coldplay and Metallica using data supplied by Spotify to uncover in the fame achieved by these 2 famous band truely is forever. 
    Focusing on their studio recordings, it explores key aspects such as album characteristics, industry trends, and the distinction between live and studio performances. The report presents statistical tables and visualizations to provide insights into the longevity and musical progression of these iconic bands.  

---


```{r setup, include=FALSE}

# Housekeeping 
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos = 'H') # set chunk defaults 
rm(list = ls()) # clear environment 
gc() # garbage collection
options(scipen = 999) #suppress scientific notation
options(dplyr.summarise.inform=F) # suppress messages 

# Load neccessary libraries 
pacman::p_load(dplyr, ggplot2, tidyverse, stringr, purrr, lubridate, gt, fmsb, hrbrthemes) ############## update

# Source all existing functions 
list.files('..//Question1/code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

# Set theme & colour palette
palette <- c("#3A7370", "#FB9072", "#FFC57B", "#CE677A", "#163940", "#2C1F2C")
theme <- set_custom_theme()

# Load data
location <- "data/Coldplay_vs_Metallica/" # set up directory path to data 
df_metallica <- read_csv(glue::glue("{location}metallica.csv")) # read in csv file
df_coldplay <- read_csv(glue::glue("{location}Coldplay.csv")) 
df_spotify <- read_rds(glue::glue("{location}Broader_Spotify_Info.rds")) # read in rds file
df_billboard_100 <- read_rds(glue::glue("{location}charts.rds"))
 
```

# Introduction 

In the dynamic world of music, artistic growth and transformation offer profound insights into the creative journeys of legendary performers. This report examines the musical evolution of Coldplay and Metallica—two iconic bands whose distinct sounds have topped charts and delighted listeners for generations. 

This analysis focuses on studio recordings to track key metrics that define their sonic identities:
* Popularity: How mainstream appeal has shifted over time
* Energy: Sense of forward motion that keeps the listener engaged - like when drums intensify and the key increases in loud music. 
* Danceability & Energy: Changes in rhythm and intensity
* Acousticness: The role of organic versus electronic elements
* Tempo: Variations in pacing and mood

Through data-driven analysis and interactive visualizations, this report reveals:
* Genre-Bending Trends: How both bands experimented beyond their roots
* Commercial vs. Artistic Balance: The interplay between chart success and creative risks
* Legacy Through Data: What makes their music endure in the streaming era

Whether you're a fan, researcher, or music enthusiast, join us in uncovering the stories behind the stats—and how these artists continue to redefine their craft.

## Context for less avid listeners: 
Metallica is a pioneering American heavy metal band known for fast tempo's, instrumentals and aggressive musicianship. Their breakthrough came with their 3rd album: 'Master of Puppets' (1986), described as one of the heaviest metal albums and named their best work.'...And Justice for All' (1988) earned their first Grammy nod, while Metallica (1991) marked a mainstream shift from their thrash metal roots, becoming their best-selling album (16M+ U.S. copies). After genre experimentation, they returned to thrash with 'Death Magnetic' (2008). Their latest release is '72 Seasons' (2023) [@wiki:metallica]. 

Coldplay, a British rock band renowned for their live performances, have sold over 100 million albums worldwide. Their debut album, 'Parachutes' (2000), featuring "Yellow," won a Brit Award and a Grammy. The follow-up, 'A Rush of Blood to the Head' (2002), repeated this success, while 'X&Y' (2005) completed what they called a trilogy. 'Viva la Vida or Death and All His Friends' (2008) also earned a Grammy, topping charts in over 30 countries. Later albums, 'Mylo Xyloto' (2011), 'Ghost Stories' (2014), 'A Head Full of Dreams' (2015), 'Everyday Life' (2019), 'Music of the Spheres' (2021), and 'Moon Music' (2024), explored genres like electronica, R&B, disco, and progressive rock [@wiki:coldplay].

```{r formatting data}

# Merge datasets
df_metallica <- df_metallica %>% 
    mutate(band = "Metallica")

df_coldplay <- df_coldplay %>% 
    mutate(band = "Coldplay")

df_metcol <- bind_rows(df_metallica, df_coldplay) %>% 
    mutate(album = coalesce(album, album_name)) %>% 
    mutate(duration = map2(duration_ms, duration, convert_duration)) #get all durations in seconds
  
# Remove song duplicates 
duplicate_words <- c("Remastered", "Remix","Rerecorded", "Re-recorded", "Deluxe", "Live", "Edit", "Edition", "Prospekt")

df_metcol_unique <- df_metcol %>%
    filter(is.na(name) | map_lgl(name, no_duplicate_words)) %>% 
    filter(is.na(album) | map_lgl(album, no_duplicate_words)) 
    
# Format dates
df_metcol_unique <- df_metcol_unique %>%
    mutate(month_year = floor_date(as.Date(release_date), "month")) 

# GOAL: Calculate days since first album release 
calculate_days_after <- function(df) {
    
  df <- df %>%
    group_by(band) %>%
    mutate(days_after = as.integer(release_date - min(release_date)))
  
  return(df)
}    
 
# GOAL: Indicate if song was performed live 
performed_live <- function(name) {
  return(str_detect(name, regex("Live", ignore_case = TRUE)))
}

```

```{r basic tables}

# Table 1 : basic summary stats
table_summary <- df_metcol_unique %>%
    filter(band %in% c("Metallica", "Coldplay")) %>%
    group_by("Band" = band) %>%
    summarise(
        "Average Popularity" = mean(popularity),
        "SD: Energy" = sd(energy),
        "SD: Danceability" = sd(danceability),
        "SD: Acousticness" = sd(acousticness),
        "SD: Tempo" = sd(tempo)
    ) %>%
    gt() %>%
    tab_header(title = "Table 1: Basic Summary Statistics for Coldplay and Metallica Albums",
             subtitle = "(SD = Standard Deviation)")

table_summary

# Best performing albums table
table_best_albums <- df_metcol_unique %>%
  filter(band %in% c("Coldplay", "Metallica")) %>%
  group_by(band, "Album" = album, "Release Date" = release_date) %>%
  summarise(
    Popularity = mean(popularity),
    Danceability = mean(danceability),
    Acousticness = mean(acousticness)
  ) %>%
  arrange(band, desc(Popularity)) %>%
  group_by(band) %>%
  top_n(3) %>% 
    gt() %>%
    tab_header(title = "Table 2: Best performing albums for Coldplay and Metallica")

table_best_albums

```

**Consistent Popularity, Divergent Paths**  
Both Coldplay and Metallica have maintained remarkable popularity throughout their careers as A-list artists. However, our analysis reveals that **Coldplay achieved higher average popularity** overall, while Metallica's success stemmed from a more consistent musical formula.  

**Metallica: The Power of Consistency**  
Metallica's music shows exceptionally low variation in key metrics:  
- Tempo (BPM)  
- Energy  
- Danceability  
- Acousticness  

The minimal standard deviations in these metrics suggest a deliberate adherence to their signature sound over time, with little experimentation across albums.  

**Coldplay: Embracing Musical Evolution**  
In contrast, Coldplay exhibits:  
- Higher variability in acousticness and energy  
- Significant deviations between songs, indicating willingness to experiment  
- Potential album-driven shifts in style  

This analysis sets the stage for deeper investigation into how these contrasting approaches have sustained two of music's most enduring acts. Might these differences be tied to specific album releases? Lets find out...  

# Comparing Characteristics

```{r stacked bar graph, fig.height=10, fig.width=8}

# Prepare data 
df_stack <- df_metcol_unique %>% 
  filter(band %in% c("Coldplay", "Metallica")) %>%
  group_by(band, month_year) %>%
  summarise(
    ave_popularity = mean(popularity, na.rm = TRUE),
    ave_energy = mean(energy, na.rm = TRUE),
    ave_danceability = mean(danceability, na.rm = TRUE),
    ave_acousticness = mean(acousticness, na.rm = TRUE),
    ave_tempo = mean(tempo, na.rm = TRUE)
  ) %>%
  pivot_longer(
    cols = starts_with("ave"),
    names_to = "category",
    values_to = "value"
  )

df_stack$month_year <- as.Date(paste(df_stack$month_year, "01", sep = "-"))

df_stack %>%
  filter(month_year > as.Date("1995-01-01")) %>% 
  ggplot(aes(x = month_year, y = value, fill = band)) +
  geom_area(position = "stack", alpha = 0.8) +
  facet_wrap(~ category, scales = "free_y", ncol = 1) +
  scale_fill_brewer(palette = "Dark2") +
  theme +
    theme(legend.position = "bottom") +
    labs(title = "Figure 1: Stacked Bar Graph of Band Characteristics Over time", x = "", y = "Count", fill = "")

```

The stacked bar graph visualization highlights key stylistic differences between Coldplay and Metallica, with distinct color coding (green for Coldplay, orange for Metallica) facilitating immediate comparison.

Coldplay's Signature Sound
* Higher acousticness: Prominent use of organic instrumentation
* Elevated danceability: Rhythmic patterns conducive to movement
* Greater popularity: Broader mainstream appeal

Metallica's Dominant Traits
* Intense energy: Characteristic high-intensity sound
* Consistent heaviness: Lower variability in energy metrics

Coldplay's fluid, evolving style contrasts sharply with Metallica's steadfast intensity, reflecting divergent artistic philosophies. This visualization underscores how these differences manifest in measurable audio features.

```{r}

table_best_albums 

```

Table 2 above depicts the most popular albums for both Coldplay and Metallica. It is interesting to note that Metallica achieves higher popularity in each of their top 3 albums (quite significantly so for the second place album) in comparison to Coldplay despite Table 1 depicting that Coldplay outscored Metallica in average popularity by quite a long shot. We also note that Coldplay's top 3 albums portray higher variance in acousticness again portraying the band's willingness to sample different music styles that win over more listeners. 

# Conclusion
This analysis reveals two distinct musical trajectories. Coldplay demonstrates greater variability in popularity and a penchant for stylistic experimentation, while Metallica maintains remarkable consistency in both appeal and sound. These contrasting approaches underscore the unique artistic identities that have sustained both bands as iconic acts. The findings illuminate how adaptability and consistency can each forge enduring musical legacies.
