---
# IMPORTANT: Change settings here, but DO NOT change the spacing AT ALL.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "QUESTION 1: Baby Names"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Talyah Greyling^[__Contributions:__  \\newline _The authors would like to thank the DNS server syncing for taking its time so that she could finish her morning cup of coffee before starting._]"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Cape Town, South Africa" # First Author's Affiliation
Email1: "23761067\\@sun.ac.za" # First Author's Email address

CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  This report is a concise analysis of baby naming trends in the US through the years utilising graphs, tables and bullet summaries to explain the findings to the client, a New York based kids’ toy design agency.
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos = 'H')

if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

list.files('code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

# Load data
Baby_names <- read_rds("data/US_Baby_names/Baby_Names_By_US_State.rds")
Top_100_billboard <- read_rds("data/US_Baby_names/charts.rds")
HBO_titles <- read_rds("data/US_Baby_names/HBO_titles.rds")
HBO_credits <- read_rds("data/US_Baby_names/HBO_credits.rds")

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

You are welcome to add some sort of an introduction here. Spend 5 minutes detailing the purpose here, e.g.

Look at the README.txt file and maybe give some background - especially also in understanding the definition of the variables.

Notice that for the below plots I created a parameters to play around with xaxis labels.

# Time time-series representation of the rank-correlation calculations 
```{r}
# Data wrangling 
top_names <- Baby_names %>% 
  top_n_names(n = 25)

# Computing Spearman Rank Correlation for a specific target year
SpearCorr <- calculate_correlations(top_names, target_year = 2000, future_years = 3)

# Apply Correlation calculation to all years
corr_results <- name_persistence(top_names, future_years = 3)

```

# Time time-series representation of the rank-correlation visualisation
```{r, warning =  FALSE, fig.align = 'center', fig.cap = "Persistence of National Top 25 Baby Names (1910-2014).\\label{Figure1}", fig.ext = 'png', fig.height = 5, fig.width = 10}

PersistPlot <- plot_name_persistence(corr_results = corr_results) 
PersistPlot

```
As you can see from Figure \ref{Figure1}, name persistence is much stronger among names for boys than names for girls. Both display trends that appear to be generally downward sloping over time. Persistence is also seen to decrease the further the future year is from the target year. Considering the client's suspicion that popular name trends have been slower to persist since the 1990s than in earlier decades their suspicion is correct for boys names, but appears to be the opposite for girls names which display with an upwards trend since the 1990s. 

\newpage 

# Popular names per decade calculations
```{r}
pop_names <- Baby_names %>% 
  top_n_names(n = 10)

decade_top_names <- top_n_names_per_decade(pop_names, top_n = 10)   
```

# Popular names per decade visualisation 
```{r, warning =  FALSE, fig.align = 'center', fig.cap = "Persistence of National Top 10 Baby Names per decade (1910-2014).\\label{Figure2}", fig.ext = 'png', fig.height = 20, fig.width = 10}

PopNamesMap <- popular_names_heatmap(decade_top_names)
PopNamesMap

```

Figure \ref{Figure2} once again shows that name persistence is much stronger among males. By displaying the information in a heatmap we can see that popularity is concentrated among less names for males than for females, contributing to their higher persistence levels.  

\newpage 

# Movie star names calculations 
```{r}
# Create a data frame of popular movies 
Popular_movies <- get_popular_movies(HBO_credits, HBO_titles)
Popular_movies 

# See if the top 10 baby names per decade are related to actors/ actresses' names from popular movies released during that decade: 

babies_after_actors <- left_join(Popular_movies, decade_top_names, by = "Name") %>% # create single data frame
    drop_na(Total_Dec_Count) %>% # filter out instances where actor name was not a popular baby name in that decade 
    filter(!(release_year > (Decade + 10))) 

top_10_female <- babies_after_actors %>%
    filter(Gender == "F") %>%
    arrange(desc(Total_Dec_Count)) %>%
    slice_head(n = 10)

top_10_male <- babies_after_actors %>%
  filter(Gender == "M") %>%               
  arrange(desc(Total_Dec_Count)) %>% 
    slice_head(n = 10)
    
top_10_female
top_10_female

```







