---
title: "The Naming Game: an Analysis on US Baby Name Data"
author: "Talyah Greyling"
date: "`r Sys.Date()`"
bibliography: Tex/ref.bib       
output:
  pagedown::html_paged:
    # template: wp_paged.html
    # css: ['wp.css', 'wp-fonts.css', 'wp-page.css']
    css: ["Template/default-fonts-Texevier.css", "Template/default-page-Texevier.css", "Template/default-Texevier.css"]
    csl: Template/harvard-stellenbosch-university.csl # referencing format used.
    template: ["Template/paged-Texevier.html"]

    toc: true 
    # change to true for a self-contained document, but it'll be a litte slower for Pandoc to render
    self_contained: TRUE
abstract: |
  This report is a concise analysis of baby naming trends in the US through the years utilising graphs, tables and bullet summaries to explain the findings to the client, a New York based kids’ toy design agency.
---

```{r setup, include=FALSE}

# Housekeeping 
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos = 'H') # set chunk defaults 
rm(list = ls()) # clear environment 
gc() # garbage collection
options(scipen = 999) #suppress scientific notation
options(dplyr.summarise.inform=F) # suppress messages 

# Load neccessary libraries 
pacman::p_load(dplyr, ggplot2, tidyverse, stringr, purrr, lubridate, gt, fmsb, hrbrthemes, knitr) ############## update 

# Source functions
list.files('code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

# Set theme & colour palette
palette <- c("#3A7370", "#FB9072", "#FFC57B", "#CE677A", "#163940", "#2C1F2C")
theme <- set_custom_theme()

# Load data
location <- "data/US_Baby_names/" # set up directory path to data 
Baby_names <- read_rds(glue::glue("{location}Baby_Names_By_US_State.rds"))
Top_100_billboard <- read_rds(glue::glue("{location}charts.rds")) 
HBO_titles <- read_rds(glue::glue("{location}HBO_titles.rds"))
HBO_credits <- read_rds(glue::glue("{location}HBO_credits.rds")) 

```

# Introduction 

This analysis examines US baby naming trends to identify which trends endure versus fade. It then explores how a cultural influence like popular movies can shape naming choices. For brevity and ease of use all findings are presented in bullet point summaries.  

```{r time-series representation of the rank-correlation calculations}

# Data wrangling 
top_names <- Baby_names %>% 
  top_n_names(n = 25)

# Computing Spearman Rank Correlation for a specific target year
SpearCorr <- calculate_correlations(top_names, target_year = 2000, future_years = 3)

# Apply Correlation calculation to all years
corr_results <- name_persistence(top_names, future_years = 3)

```

\newpage
# Persistence of baby names 
```{r, warning =  FALSE, fig.align = 'center', fig.cap = "Persistence of National Top 25 Baby Names (1910-2014) by Spearman rank-correlation", fig.ext = 'png', fig.height = 5, fig.width = 7}

PersistPlot <- plot_name_persistence(corr_results = corr_results) 
PersistPlot

```
Insights from Figure 1 (above):
* Name persistence is much stronger among boys than girls.
* Generally boy and girl names display downward sloping persistence trends over time.
* Persistence becomes weaker for boys and girls the further the future year is from the target year. 
* The client's suspicion on slower to persistence since the 1990s is: 
    * Correct for boys names.
    * Incorrect for girls names which display with an unexpected upwards trend since the 1990s.  

\newpage 
# Popular Names 
```{r Popular names per decade calculations}
pop_names <- Baby_names %>% 
  top_n_names(n = 10)

decade_top_names <- top_n_names_per_decade(pop_names, top_n = 10)   
```

```{r, warning =  FALSE, fig.align = 'center', fig.cap = "Most Popular Baby Names per decade (1910-2014)", fig.ext = 'png', fig.height = 10, fig.width = 7}

PopNamesMap <- popular_names_heatmap(decade_top_names)
PopNamesMap

```
Insights from Figure 2 (above):
* Name persistence is much stronger among males. 
* Popularity is concentrated among less names for males, contributing to their higher persistence levels.  
* Most persistently popular name over all decades:
    * Girls: Mary
    * Boys: William 
* High popularity before the 1960s:
    * Girls: Nancy, Margaret, Helen, Patricia, Barbara, Mary
    * Boys: Thomas, Charles, Richard, William, John, Robert, James
* High popularity after the 1960s:
    * Girls: Elizabeth, Amanda, Ashley, Emily 
    * Boys: Daniel, Matthew, Joshua, Christopher, James, Michael 

\newpage 
# Movie star names calculations 
```{r}
# Create a data frame of popular movies 
Popular_movies <- get_popular_movies(HBO_credits, HBO_titles)
Popular_movies 

# See if the top 10 baby names per decade are related to actors/ actresses' names from popular movies released during that decade: 

babies_after_actors <- left_join(Popular_movies, decade_top_names, by = "Name") %>% # create single data frame
    drop_na(Total_Dec_Count) %>% # filter out instances where actor name was not a popular baby name in that decade 
    filter(!(release_year > (Decade + 10))) 

top_10_female <- babies_after_actors %>%
    filter(Gender == "F") %>%
    arrange(desc(Total_Dec_Count)) %>%
    slice_head(n = 10)

top_10_male <- babies_after_actors %>%
  filter(Gender == "M") %>%               
  arrange(desc(Total_Dec_Count)) %>% 
    slice_head(n = 10)

female_actors_table <- kable(top_10_female, 
      format = "html",  
      caption = "Top 10 Female Names Coinciding With Actor Names")
female_actors_table

male_actors_table <- kable(top_10_male, 
      format = "html",  
      caption = "Top 10 Male Names Coinciding With Actor Names")
male_actors_table

```

# Do popular movies influence popular baby names? 
## Insights from Tables above:
* TMDB = The Movie Database: The percentage score for a movie or TV show is by database users on a 10-star scale. 
* This analysis was less informative since the most popular baby names, Mary for females and James for males, that coincide with actor names were shown already be popular prior to the movie release dates when looking at the heat map.  







