---
title: "Streaming Service Success: an Analysis on Netflix Data"
author: "Talyah Greyling"
date: "`r Sys.Date()`"
bibliography: Tex/ref.bib       
output:
  pagedown::html_paged:
    # template: wp_paged.html
    # css: ['wp.css', 'wp-fonts.css', 'wp-page.css']
    css: ["Template/default-fonts-Texevier.css", "Template/default-page-Texevier.css", "Template/default-Texevier.css"]
    csl: Template/harvard-stellenbosch-university.csl # referencing format used.
    template: ["Template/paged-Texevier.html"]

    toc: false
    # change to true for a self-contained document, but it'll be a litte slower for Pandoc to render
    self_contained: TRUE
    
abstract: |
    This html document contains an analysis of Netflix data to uncover the dos and don'ts of streaming services. Key statistics, graphs and tables are accompanied by short summaries ideal for any busy investor to make quick & efficient investment calls. 

---

```{r setup, include = FALSE}

# Housekeeping 
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos = 'H') # set chunk defaults 
rm(list = ls()) # clear environment 
gc() # garbage collection
options(scipen = 999) #suppress scientific notation
options(dplyr.summarise.inform=F) # suppress messages 

# Load neccessary libraries 
pacman::p_load(dplyr, ggplot2, tidyverse, stringr, purrr, lubridate, gt, fmsb, hrbrthemes) ############## update 

# Source all existing functions 
list.files('..//Question1/code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('..//Question2/code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

# Set theme & colour palette
palette <- c("#3A7370", "#FB9072", "#FFC57B", "#CE677A", "#163940", "#2C1F2C")
theme <- set_custom_theme()

# Load data 
location <- "data/netflix/"
df_titles <- read_rds(glue::glue("{location}titles.rds"))
df_credits <- read_rds(glue::glue("{location}credits.rds"))
df_movies <-  read_csv(glue::glue("{location}netflix_movies.csv"))

```

\newpage 
# Introduction
Recent declines in Netflix subscribers and stock value may be signaling shifting industry dynamics. Typically drama, comedy, and action genres drive the highest viewer engagement [@ContentEngine2023], offering critical insights for new streaming platforms. This analysis summarises key trends in successful content strategies to guide potential investors in the competitive streaming market.

# Informative Graphs 
```{r genre plot wrangling}

# Prepare data for genre distribution plot
df_titles <- add_main_genre(df_titles)
genre_distribution <- counts_genre(df_titles)

# Prepare data for top n genres plot
top_10_genres <- get_top_n_genres(df_titles, 10)

```

```{r, fig.align = 'center', fig.cap = "Distribution of genres on Netflix", fig.ext = 'png', fig.height = 5, fig.width = 7}

# Plot genre distribution
genre_distribution_plot <- draw_genre_distribution(genre_distribution)
genre_distribution_plot

```

```{r, fig.align = 'center', fig.cap = "Top genres according to audience", fig.ext = 'png', fig.height = 5, fig.width = 7}

# Plot top 10 genres
top_n_genres_plot <- draw_top_n_genres(top_10_genres)
top_n_genres_plot

```

```{r duration plot wrangling}

# Create duration data frame
df_duration <- left_join(df_titles, df_movies, by = "title") %>% 
    filter(!is.na(imdb_score)) %>%
    mutate(minutes = sub(" min", "", duration))

# Prepare data for duration distribution plot 
duration_distribution <- get_duration_distribution(df_duration)

# Prepare data for top n durations by average audience score plot
top_10_durations <- get_top_n_durations(df_duration, 10)

```

```{r, fig.align = 'center', fig.cap = "Distribution of durations on Netflix", fig.ext = 'png', fig.height = 5, fig.width = 7}

# Plot duration distribution
duration_distribution_plot <- draw_duration_distribution(duration_distribution)
duration_distribution_plot

```

```{r, fig.align = 'center', fig.cap = "Top genres according to audience", fig.ext = 'png', fig.height = 5, fig.width = 7}

# Plot top 10 durations by average audience score
top_n_durations_plot <- draw_top_n_durations(top_10_durations)
top_n_durations_plot

```

\newpage
# Summary of statistics 
The data reveals a clear trend: drama and comedy films make up the bulk of Netflix’s catalog, while animation and fantasy titles are significantly underrepresented. Yet, despite their scarcity, the fantasy, animation, and drama genres consistently earn high IMDb ratings, suggesting a strong audience appeal.

To better align with viewer demand, Netflix could benefit from boosting its fantasy and animation offerings, leveraging their proven popularity to attract and retain subscribers. At the same time, maintaining a robust selection of drama and comedy—already well-established favorites—would help sustain its current audience.

A balanced content strategy, combining expanded investment in high-performing niche genres (like fantasy and animation) with steady support for mainstream hits (like drama and comedy), could enhance Netflix’s competitive edge and broaden its subscriber base.

\newpage 
# References
