---
title: "Booming Billions: an Analysis on Global Wealth Data"
author: "Talyah Greyling"
date: "`r Sys.Date()`"
bibliography: Tex/ref.bib       
output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 
    fig_height: 3.5
    toc: true 
    self_contained: TRUE
    
abstract: |
    This pdf document contains an analysis on a collection of global billionaires using data supplied by Forbes to uncover the underlying patterns to becoming one of the world's wealthiest. Do developing countries deliver more self-made men? Is the software industry the key to a comfortable life? Read this paper to find out. 
    
---

```{r setup, include = FALSE}

# Housekeeping 
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos = 'H') # set chunk defaults 
rm(list = ls()) # clear environment 
options(scipen = 999) #suppress scientific notation
options(dplyr.summarise.inform=F) # suppress messages 

# Load neccessary libraries 
pacman::p_load(dplyr, ggplot2, tidyverse, stringr, purrr, lubridate, gt) ############## update 

# Source all existing functions 
list.files('..//Question1/code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('..//Question2/code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('..//Question3/code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

# Set theme & colour palette
palette <- c("#3A7370", "#FB9072", "#FFC57B", "#CE677A", "#163940", "#2C1F2C")
theme <- set_custom_theme()

# Load data
df_billions <- "data/Billions/billionaires.csv" %>% read_csv_with_col_type()

```

# Introduction

Global wealth inequality has reached unprecedented levels, with just 8% of the world’s population owning 85% of its wealth and most of this being concentrated in the top 1% [@Piketty2015]. Against this backdrop, Forbes surveyed South Africa’s wealthiest individuals and identified a pressing need to analyze shifting wealth patterns across three decades—from the 1990s to the mid-2010s. With disparities widening both within and between nations [@Mittelman2007], this research aims to uncover critical insights that could inform future economic strategies. If successful, Forbes plans to expand this database, offering a vital tool to address one of today’s most urgent challenges: an increasingly unequal world.

## Exploration Goals
Two main claims (made by one of Forbs' participants) will be explored in this document: 
(1) "In the US, you saw an increasing number of new billionaires emerge that had little to no familial ties to generational wealth. Other developed markets and emerging markets tend to have less entrepreneurial successes and tend to house mostly inherited wealth."
(2) “Most new self-made millionaires are in software, compared to consumer services type industries in the 90s. This is related to different countries’ GDP, of course, with richer countries providing more innovation in consumer services."

# Exploring claim number 1 
* This claim entails comparing the prevalence of self-made billionaires in the US versus inherited wealth dominance in other developed and emerging markets.

```{r first claim}

# Prepare data for first claim plot 
df_us_billions <- counts_new_us(df_billions)
df_other_billions <- counts_new_other(df_billions)

"Counts of new vs inhereted wealth in the US per decade"
us_plot_first_claim <- draw_first_claim(df_us_billions, input_title = "Counts of new vs inhereted wealth in the US per decade")
us_plot_first_claim

ggplot(df_us_billions, aes(fill = new_status, y = count, x = year)) + 
    geom_bar(position="dodge", stat="identity")


# Create first claim plot (grouped bar chart)
first_claim_plot <- bind_rows(df_us_billions, df_other_billions) %>% 
    ggplot(aes(fill=condition, y=value, x=specie)) + 
    geom_bar(position="dodge", stat="identity")

```

# Exploring claim number 2
* Analyze the shift from consumer services (1990s) to software as the primary industry for self-made millionaires, correlating with national GDP levels.


# Additional insights 
* 


