---
title: "Booming Billions: an Analysis on Global Wealth Data"
author: "Talyah Greyling"
date: "`r Sys.Date()`"
bibliography: Tex/ref.bib       
output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 
    fig_height: 3.5
    toc: true 
    self_contained: TRUE
    
abstract: |
    This pdf document contains an analysis on a collection of global billionaires using data supplied by Forbes to uncover the underlying patterns to becoming one of the world's wealthiest. Do developing countries deliver more self-made men? Is the software industry the key to a comfortable life? Read this paper to find out. 
    
---

```{r setup, include = FALSE}

# Housekeeping 
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos = 'H') # set chunk defaults 
rm(list = ls()) # clear environment 
options(scipen = 999) #suppress scientific notation
options(dplyr.summarise.inform=F) # suppress messages 

# Load neccessary libraries 
pacman::p_load(dplyr, ggplot2, tidyverse, stringr, purrr, lubridate, gt) ############## update 

# Source all existing functions 
list.files('..//Question1/code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('..//Question2/code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('..//Question3/code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

# Set theme & colour palette
palette <- c("#3A7370", "#FB9072", "#FFC57B", "#CE677A", "#163940", "#2C1F2C")
theme <- set_custom_theme()

# Load data
df_billions <- "data/Billions/billionaires.csv" %>% read_csv_with_col_type()

```

# Introduction

Global wealth inequality has reached unprecedented levels, with just 8% of the world’s population owning 85% of its wealth and most of this being concentrated in the top 1% [@Piketty2015]. Against this backdrop, Forbes surveyed South Africa’s wealthiest individuals and identified a pressing need to analyze shifting wealth patterns across three decades—from the 1990s to the mid-2010s. With disparities widening both within and between nations [@Mittelman2007], this research aims to uncover critical insights that could inform future economic strategies. If successful, Forbes plans to expand this database, offering a vital tool to address one of today’s most urgent challenges: an increasingly unequal world.

## Exploration Goals
The two main research questions that will be discussed in this document are as follows: 
* Compare the prevalence of self-made billionaires in the US versus inherited wealth dominance in other developed and emerging markets.
* Analyze the shift from consumer services (1990s) to software as the primary industry for self-made millionaires, correlating with national GDP levels.

```{r}

```





